The Lasting Shadows of Economic Downturns

### 2024-02-14 outputs/paper/paper.qmd
Update paper.qmd

# Introduction
*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here](update the link!!!)***

# Data
This paper replicates and extends the analyses from the original study by using a range of datasets obtained from various publicly accessible sources. Stuart has categorized the datasets into distinct files, reflecting their respective roles in the analysis.

## Sources
```{r}
### few rows for unemployment or usa_personal_income data presented here as example ###
```
We used data from two sources for our research. First, we utilized data from the Bureau of Economic Analysis (BEA) Regional Economic Accounts[@BEA]. These datasets were downloaded in March 2015 from [https://apps.bea.gov/regional/downloadzip.cfm](https://apps.bea.gov/regional/downloadzip.cfm) by Stuart. The BEA occasionally revises the methodology used to generate these data, and accordingly, updates all files to reflect these changes. As a result, data downloaded on subsequent dates might present some variations. The datasets of particular interest included 'CA1: Personal Income Summary: Personal Income, Population, Per Capita Personal Income', which covers Personal Income, Population, and Per Capita Personal Income, and 'CA25: Total Full-Time and Part-Time Employment by SIC Industry', offering detailed insights into income components and industry-specific earnings.

@@

# Results
```{r}
### the line graphs of average unemployment rate by four most populated states###
### which is something like simulated plot1###
```

@@

# Discussion
## How Recession Impacts on Unemployment
## How Recession Impacts on Personal Income
## How Recession Changes the Business Patterns
## Limitations and Next Steps
Due to limitations in accessing relevant datasets, this paper will not explore the diverse social impacts of recessions, specifically on birth rates and education, which were initially of primary interest. Drawing from Stuart's foundational research, we quantitatively assess the prolonged impacts of the 1980–1982 recession on individuals who were in their formative years—children, adolescents, and young adults—at the onset of this economic downturn. Stuart's findings reveal significant long-term detriments to educational attainment and income levels. Specifically, for individuals aged 0–10 in 1979, a 10 percent reduction in per capita earnings from 1979 to 1982 within their birth county correlates with a 4.7 percentage point (14.6 percent) reduction in the attainment of four-year college degrees, a decrease in earned income by $2,100 (5.2 percent), and a 1.6 percentage point (13.0 percent) increase in the likelihood of living in poverty during 2000–2013. The adverse effects on college graduation rates are notably pronounced and consistently significant for individuals aged 0–13 in 1979, and diminish to small and statistically non-significant levels for those aged 14–22. This pattern indicates that the recession's impact likely stems from a reduction in childhood human capital or a sustained decrease in parental financial capacity to support higher education. Future research will address the effects on education and birth rates once relevant and accessible datasets become available.

\newpage
# References

>>>>

```{r}
#| message: false
#| echo: false
#| warning: false
#### Workspace set up ####
# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("here")
# install.packages("janitor")
# install.packages("kableExtra")

library(tidyverse) # data-related packages
library(knitr) # make tables
library(here) # manage file paths
library(janitor) # clean data
library(kableExtra) # table-related packages

#### Read in cleaned data ####
cleaned_unemployment <- read_csv(file = here("outputs/data/cleaned_unemployment.csv"), show_col_types = FALSE)
cleaned_four_states <- read_csv(file = here("outputs/data/cleaned_four_states.csv"), show_col_types = FALSE)
cleaned_personal_income <- read_csv(file = here("outputs/data/cleaned_personal_income.csv"), show_col_types = FALSE)
cleaned_bussiness_pattern1 <- read_csv(file = here("outputs/data/cleaned_bussiness_pattern1.csv"), show_col_types = FALSE)
cleaned_bussiness_pattern2 <- read_csv(file = here("outputs/data/cleaned_bussiness_pattern2.csv"), show_col_types = FALSE)
```

# Introduction

*This reproduction was performed after a replication on the Social Science Reproduction platform: [**link here**](update%20the%20link!!!)*

# Data

This paper replicates and extends the analyses from the original study by using a range of datasets obtained from various publicly accessible sources. Stuart has categorized the datasets into distinct files, reflecting their respective roles in the analysis.

## Sources

```{r}
### few rows for unemployment or usa_personal_income data presented here as example ###
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-1
#| tbl-cap: Sample 1
kable(cleaned_four_states, format = "latex", booktabs = TRUE) |>
  kable_styling(latex_options = c("striped", "scale_down"), font_size = 7)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-2
#| tbl-cap: Sample 2
kable(head(cleaned_unemployment, 30), format = "latex", booktabs = TRUE) |>
  kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-3
#| tbl-cap: Sample 3
kable(cleaned_personal_income, format = "latex", booktabs = TRUE) |>
  kable_styling(latex_options = c("striped", "scale_down"))
```

We used data from two sources for our research. First, we utilized data from the Bureau of Economic Analysis (BEA) Regional Economic Accounts[@BEA]. These datasets were downloaded in March 2015 from <https://apps.bea.gov/regional/downloadzip.cfm> by Stuart. The BEA occasionally revises the methodology used to generate these data, and accordingly, updates all files to reflect these changes. As a result, data downloaded on subsequent dates might present some variations. The datasets of particular interest included 'CA1: Personal Income Summary: Personal Income, Population, Per Capita Personal Income', which covers Personal Income, Population, and Per Capita Personal Income, and 'CA25: Total Full-Time and Part-Time Employment by SIC Industry', offering detailed insights into income components and industry-specific earnings.

@@

# Results

```{r}
### the line graphs of average unemployment rate by four most populated states###
### which is something like simulated plot1###
```

@@

# Discussion

## How Recession Impacts on Unemployment

## How Recession Impacts on Personal Income

## How Recession Changes the Business Patterns

## Limitations and Next Steps

Due to limitations in accessing relevant datasets, this paper will not explore the diverse social impacts of recessions, specifically on birth rates and education, which were initially of primary interest. Drawing from Stuart's foundational research, we quantitatively assess the prolonged impacts of the 1980–1982 recession on individuals who were in their formative years—children, adolescents, and young adults—at the onset of this economic downturn. Stuart's findings reveal significant long-term detriments to educational attainment and income levels. Specifically, for individuals aged 0–10 in 1979, a 10 percent reduction in per capita earnings from 1979 to 1982 within their birth county correlates with a 4.7 percentage point (14.6 percent) reduction in the attainment of four-year college degrees, a decrease in earned income by \$2,100 (5.2 percent), and a 1.6 percentage point (13.0 percent) increase in the likelihood of living in poverty during 2000–2013. The adverse effects on college graduation rates are notably pronounced and consistently significant for individuals aged 0–13 in 1979, and diminish to small and statistically non-significant levels for those aged 14–22. This pattern indicates that the recession's impact likely stems from a reduction in childhood human capital or a sustained decrease in parental financial capacity to support higher education. Future research will address the effects on education and birth rates once relevant and accessible datasets become available.

\newpage

# References
